---
title: "Netwok Analysis Project"
author: "Eloy Pérez Gómez, Álvaro Novillo Correas & Jorge Sánchez Polo"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(jsonlite)
library(igraph)
```


**1) Explain the network including which vertices and edges form it, as well as any additional attributes that may form the network, its vertices and/or its edges.**

We begin by loading the network

```{r}
# Set the directory path where the JSON files are located
directory_path <- "C:/Users/ljspv/Downloads/Network_Analysis_Kaggle_Star_Wars-main/Network_Analysis_Kaggle_Star_Wars-main/data/" #CHANGE IT WITH YOUR DATA PATH

# Get the list of all JSON files in the directory
json_files <- list.files(path = directory_path, pattern = "*.json", full.names = TRUE)

# Display the list of JSON files
print(json_files)

# Choose the desired JSON file
desired_file <- json_files[22]

# Load the JSON file into R
json_data <- fromJSON(desired_file)
```

Access the nodes and edges
```{r}
nodes <- json_data$nodes
edges <- json_data$links
```

**2) Describe what the analysis of such a network may be of interest. That is, what are the points that make the network of interest to a general audience.**



**3) Make a first graph of the network and describe the main characteristics including whether the network is directed or undirected, if there are subnetworks that may be of interest, if there are loops and/or multi-edges, etc.**

Create a graph from the data
```{r}
network <- graph_from_data_frame(edges, directed = T)
E(network)$weight <- edges$value
```


```{r}
# Set the color of each node
V(network)$color <- nodes$colour

# Set the label of each node
V(network)$label <- nodes$name
```

Plot the graph
```{r}
A = as_adjacency_matrix(network, attr="weight")
plot(network)
```


**4) Compute the order and size of the network and certain induced subnetworks of interest and draw conclusions about these values. If the network is directed, repeat the operation for in, out and total.**

**5) Obtain the degree distribution with and without weights (if any) and comment on the results obtained. Again, if the network is directed, repeat the operation for in, out and total.**

**6) Obtain the number of components of the network (weak and strong if the network is directed).**

**7) Compute the diameter with and without weights (if any) and determine the farthest vertices along with the path joining them. Take into account whether the network is directed or not.**

**8) Visualize the network. Obtain the graphs that appear on the slides and the code and, very importantly, try to obtain the best possible visualization by using colors and varying all aspects of the graph necessary to do so.**

##Davidson-Harel layout 
```{r}
plot(subnetwork1, layout = layout_with_dh, edge.color = 'darkgrey',layout = m2,edge.width = (edges$value - min(edges$value))/(max(edges$value) - min(edges$value))*(7-2) + 2 , vertex.size=(nodes$value - min(nodes$value))/(max(nodes$value) - min(nodes$value))*(20-5) + 7)
```

##Fruchtermann and Reingold layout
```{r}
plot(subnetwork1, layout = layout_with_fr, edge.color = 'darkgrey',layout = m2,edge.width = (edges$value - min(edges$value))/(max(edges$value) - min(edges$value))*(7-2) + 2 , vertex.size=(nodes$value - min(nodes$value))/(max(nodes$value) - min(nodes$value))*(20-5) + 7)
```

##GEM force-directed layout
```{r}
plot(subnetwork1, layout = layout_with_gem, edge.color = 'darkgrey',layout = m2,edge.width = (edges$value - min(edges$value))/(max(edges$value) - min(edges$value))*(7-2) + 2 , vertex.size=(nodes$value - min(nodes$value))/(max(nodes$value) - min(nodes$value))*(20-5) + 7)
```

##Graphopt layout
```{r}
plot(subnetwork1, layout = layout_with_graphopt, edge.color = 'darkgrey',layout = m2,edge.width = (edges$value - min(edges$value))/(max(edges$value) - min(edges$value))*(7-2) + 2 , vertex.size=(nodes$value - min(nodes$value))/(max(nodes$value) - min(nodes$value))*(20-5) + 7)
```

##Kamada and Kawai layout
```{r}
plot(subnetwork1, layout = layout_with_kk, edge.color = 'darkgrey',layout = m2,edge.width = (edges$value - min(edges$value))/(max(edges$value) - min(edges$value))*(7-2) + 2 , vertex.size=(nodes$value - min(nodes$value))/(max(nodes$value) - min(nodes$value))*(20-5) + 7)
```

##Multidimensional Scaling layout
```{r}
plot(subnetwork1, layout = layout_with_mds, edge.color = 'darkgrey',layout = m2,edge.width = (edges$value - min(edges$value))/(max(edges$value) - min(edges$value))*(7-2) + 2 , vertex.size=(nodes$value - min(nodes$value))/(max(nodes$value) - min(nodes$value))*(20-5) + 7)
```
